---
title: "Project 2 notebook"
output: html_notebook
---

# Load packages
```{r}
library(tidyverse)
library(ggplot2)
```

# Read in Data (takes time)
```{r}
NEI <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")

# NEI year stored as int. We'll set to factors
NEI$year <- as.factor(NEI$year)

# NEI type stored chr. We'll set to factors
NEI$type <- as.factor(NEI$type)

```
# Question 1

Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Using the base plotting system, make a plot showing the total PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.
```{r}
# Plot1.R
totals <- NEI %>%
  group_by(year) %>%
  summarise(Total = sum(Emissions, na.rm=TRUE)/1e6)

# png device for plotting
png(filename = "plot1.png", width = 480, height = 480)

barplot(totals$Total, names.arg = totals$year,
        col = "#fa9fb5",
        main = "Total PM2.5 Emissions\n in United States",
        ylab = "Amount (10^6 tons)")

dev.off() # close device
```

# Question 2

Have total emissions from PM2.5 decreased in the Baltimore City, Maryland fips == "24510" from 1999 to 2008? Use the base plotting system to make a plot answering this question.
```{r}
# Plot2.R

# Baltimore City, Maryland: fips = 24510
bc_m <- NEI %>%
  filter(fips == "24510") %>%
  group_by(year) %>%
  summarise(Total = sum(Emissions, na.rm = TRUE))


# png device for plotting
#png(filename = "plot2.png", width = 480, height = 480)

barplot(bc_m$Total, names.arg = bc_m$year,
        col = "#9ecae1",
        main = "Total PM2.5 Emissions\n in Baltimore City, Maryland",
        ylab = "Amount (tons)")

#dev.off() # close device
```

# Question 3

Of the four types of sources indicated by the type (point, nonpoint, onroad, nonroad) variable, which of these four sources have seen decreases in emissions from 1999 - 2008 for Baltimore City? Which have seen increases in emissions from 1999 - 2008? Use the ggplot2 plotting system to make a plot answer this question.
```{r}
# Plot3.R

# Baltimore City, Maryland: fips = 24510
bc_m_type <- NEI %>%
  filter(fips == "24510") %>%
  group_by(type, year) %>%
  summarise(Total = sum(Emissions, na.rm = TRUE))

# png device for plotting
png(filename = "plot3.png", width = 480, height = 480)

ggplot(bc_m_type, aes(x = year, y = Total, fill = type)) + 
  geom_col(position = "dodge", colour = "black") +
  ylab("Total (tons)") + xlab(NULL) +
  ggtitle("Total PM2.5 Emissions\n in Baltimore City, Maryland")


dev.off() # close device
```

# Question 4

Across the United States, how have emissions from coal combustion-related sources changed from 1999 - 2008?
```{r}
```

# Question 5

How have emissions from motor vehicle sources changed from 1999 - 2008 in Baltimore City?
```{r}
```

# Question 6

Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in Los Angeles County, California fips == "06037". Which city has seen greater changes over time in motor vehicle emissions?
```{r}
```
